---
title: "ENVS-193DS_homework-03"
author: "Isaac Miller"
date: "2026-03-05"
format: 
  html:
    toc: true # use this to display a table of contents
editor: 
  markdown: 
    wrap: 72
---

# Set Up

[Link to my GitHub
Repository](https://isaacmiller1113.github.io/ENVS-193DS_homework-03/code/ENVS-193DS_homework-03.html)

```{r}
#| label: set up chunk 
#| message: false

# Load in packages
library(tidyverse)     # general use
library(here)          # managing file paths relative to project root
library(janitor)       # cleaning data frames
library(readxl)        # importing data from excel files (.xls or .xlsx)
library(ggeffects)     # generating model predictions
library(gtsummary)     # generating summary tables for models
library(MuMIn)         # model selection and model inference
library(broom.helpers) # tidying and labeling model outputs

# Read in data objects
salinity <- read_csv(here("data", "salinity-pickleweed.csv")) |> 
  clean_names() |>  # Store salinity data
  rename(salinity_mscm = salinity_m_s_cm)

# insert code to read in my data and label it as an object called "my_data"
my_data <- read_csv(here("data", "my_data_hw3.csv"))

```

# Read in my personal data



# Problem 1. Slough soil salinity

You are working at a restoration site where you are managing planting of
California pickleweed (Salicornia virginica) along a brackish slough
(i.e. there is a mixture of fresh water and salt water).

You decide to measure plant growth for individual pickleweed plants by
plucking an individual out of the ground and measuring the biomass (in
g). You also measure soil salinity (as electrical conductivity in units
of millisiemens per centimeter, or (mS/cm) at the location in which the
individual was growing. Admittedly, this isn’t a perfect study, but it’s
what you can do with the time and resources you have!

## a. An appropriate test

To determine the strength of the relationship between soil salinity
(mScm) and California pickleweed biomass (g), the two appropriate tests
are Pearson’s ($r$) correlation and Spearman’s ($\rho$) rank
correlation. Pearson's correlation is a parametric test that measure the
strength and direction of the relationship between variables (biomass
and salinity). Spearmans rank correlation is a non-parametric version
that does not depend on the assumption that the data is normally
distributed or continuous, and measures the monotonic relationship; the
data is ranked from smallest to largest and correlation is calculated
based on the ranked numbers rather than the actual measurements.

## b. Create a visualization

Create a visualization that would be appropriate for showing the
relationship between soil salinity (in mS/cm) and California pickleweed
biomass (in g).

```{r}
#| label: data visulaization

ggplot(data = salinity,
       aes(x = salinity_mscm,
           y = pickleweed)) +
  geom_point(color = "darkgreen",
            alpha = 0.6) + 
  labs(x = "Soil salinity (mS/cm)",
       y = "Pickleweed biomass (g)",
       title = "As Salinity Increases, Pickleweed Biomass also Increases") + 
  theme_bw()


```

## c. Check your assumptions and run your test

### 1. check your assumptions

```{r}
#| label: create model object to fit linear model
# soil salinity (mScn) as a function of plant biomass (g)

salinity_model <- lm( # model object using linear model
  pickleweed ~ salinity_mscm, # formula (response ~ predictor)
  data = salinity # data frame 
)

```

```{r}
#| label: diagnostic plots

# base R residuals 
par(mfrow = c(2, 2)) # display plots in 2x2 grid
plot(salinity_model) # plotting residuals of model object 

```

The data shows that the residualas have a constant variance
(homoscedastic) accross the range of fitted valies according to the
Residuals vs Fitted and Scale-Location plots as they are evenly and
randomly distributed. The residuals (errors) seem to be normally
distributed according to the QQ Residuals plot as the data more or less
follows a straight line. According to the Residuals vs Leverage plot
there are no major outliers that might influence model estimates of
slope or y-intercept.

### run your test

```{r}
 
# more information about the model coefficients (slope and intercept)
summary(salinity_model)
```

For each one unit increase in salinity (mScm) there is an increase in
biomass of 2.08 + 0.72 grams

```{r}
#| label: visualize model predictions

salinity_preds <- ggpredict( # create prediction object 
  salinity_model, # model object
  terms = "salinity_mscm")  # predictor variable 

# display the predictions
print(salinity_preds, n = Inf) 
```

```{r}
#| label: pearson's r correlation test

cor.test(salinity$pickleweed, salinity$salinity_mscm,
         method = "pearson")

```

There is a moderate relationship between soil salinity (mS/cm) and pickleweed biomass (g) (Pearson's r = 0.53, t(21) = 2.90, p > 0.001, $\alpha$ = 0.05)

linear relationship (visual) variables are continuous (qq plot) normally distributed indeopendent observations

Additionally, write 1-3 sentences describing:

The data shows that the residualas have a constant variance
(homoscedastic) accross the range of fitted valies according to the
Residuals vs Fitted and Scale-Location plots as they are evenly and
randomly distributed. The residuals (errors) seem to be normally
distributed according to the QQ Residuals plot as the data more or less
follows a straight line. According to the Residuals vs Leverage plot
there are no major outliers that might influence model estimates of
slope or y-intercept.

I checked the assumptions of linearity, homoscedasticity (equal variance), and normality of residuals by visually inspecting the Residuals vs Fitted, Scale-Location, and Normal Q-Q plots generated from the model object. My assessment is that the assumptions are met because the residuals are randomly distributed without a "fan" shape and the Q-Q plot points closely follow the 1:1 diagonal line.

which assumptions you checked how you checked your assumptions your
assessment of your assumption checks In the section in which you run
your test, be sure your code is annotated and the outputs are visible.

COMMIT

## d. Results communication

To evaluate the strength of the relationship between pickleweed biomass and soil salinity, I used a pearson's r corrleation test. 

We found a moderate positive relationship between soil salinity (mS/cm) and pickleweed biomass (g) (Pearson's r = 0.53, t(21) = 2.90, p > 0.001, $\alpha$ = 0.05)

From our model coefficient estimates, we found that for each one unit
increase in salinity (mScm) there is an increase in biomass of 2.08 ±
0.72 grams.

COMMIT

## e. Test implications

You’re working on a team of people at this restoration site who are also
concerned about pickleweed planting. In 2-3 sentences, write what you
would communicate to them about the results of this test and what it
means for pickleweed planting success at your site.

Be cognizant of your audience as you are writing: what would they need
to know to take action?

COMMIT

## f. Double check your own work.

# Problem 2. Personal data

## a. Updating your visualizations

## b. Captions

# Problem 3. Affective visualization

In this problem, you will create an affective visualization using your personal data in preparation for workshops during weeks 9 and 10.

In lecture, we talked about the three vertices of data visualization: 1) exploratory, 2) affective, and 3) communicative. We’ve done a lot of exploratory and communicative visualization, but have yet to think about affective visualization.

When thinking of affective visualization, you can expand your ideas of what data visualization could be. Some examples of affective visualizations include:

Jill Pelto’s paintings
Lorraine Woodruff-Long’s warming strips quilt
Stefanie Posavec and Giorgia Lupi’s Dear Data project
Environmental Graphiti
See slides from the data visualization lecture for more ideas.

## a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences).



## b. Create a sketch (on paper) of your idea.

## c. Make a draft of your visualization.

## d. Write an artist statement.

## e. Prep your materials to share in class

# Problem 4. Statistical critique

## a. Revisit and summarize

## b. Visual clarity

## c. Aesthetic clarity

## d. Recommendations
